\section{Introduction} \label{intro}

Sleep is considered to be a ubiquitous and necessary behaviour amongst
animals\cite{siegel_all_2008,cirelli_is_2008}.
However, its real physiological functions remain debated.
In vertebrates, electrophysiological recordings (\gls{eeg}), and in particular recordings of the global electrical activity in the brain,
and electromyogram recordings (\gls{emg}) of muscular activity,
have been extensively used in the last
century to study the structure of sleep \cite{loomis_distribution_1938,aserinsky_regularly_1953}.
They have the advantage of being non-invasive and relatively high throughput.
Today, \gls{eeg} remains one of the main assets in the study of sleep physiology.

Rodents, in particular mice and rats, have proven to be very successful models to
understand the mechanisms of sleep in mammals\cite{toth_animal_2013}.
Classically, there are three distinct main types of sleep-related behaviours: wakefulness, \gls{nrem} sleep and \gls{rem} 
sleep which are collectively referred to as \emph{vigilance states}.
Vigilance states are usually defined on the basis of \gls{eeg} and \gls{emg} (fig.~\ref{fig:sleep_description}).
When awake (WAKE), animals tend to have a high muscular activity which translates to high amplitude voltage changes in the \gls{emg}\cite{crisler_sleep-stage_2008}.
During wakefulness, \gls{eeg} is dominated by relatively low amplitude oscillations of frequencies
between six and ten hertz often referred to as theta waves. 
In contrast, \gls{nrem} sleep, also called slow wave sleep, is a period of muscular inactivity (low \gls{emg})
dominated by slow oscillations (below 4Hz) of high amplitude named delta waves.
\cite{vyazovskiy_nrem_2014}
The third state, \gls{rem} sleep, is characterised by a complete lack of muscular activity (atony) and an \gls{eeg} activity very similar to the awake state.
\gls{rem} sleep is the least prevalent of all three stages, and represents generally  20\% of all sleeping time.
The prevalence of these three states as well as their temporal organisation are extremely important observations in sleep research
\cite{mieda_orexin_2004,pang_unexpected_2009}

\input{./figures/sleep_description}


Although definitions of sleep stages appear straightforward, many cases are ambiguous in practice.
For instance, it is difficult to characterise transitions between two states.
In addition, there are many sources of variability including how surgery was performed, the type of instrument used and inter-animal variability.
The quality of the acquisition can also be made considerably worse by noisy signals or by the presence of artefacts.
For these reasons, sleep scoring, \ie{} the attribution of discrete vigilance states to electrophysiological time series,
is traditionally performed by trained human experts.
Such manual annotation is very time consuming; several hours of work have been reported in order to score 24h of recording\cite{sunagawa_faster:_2013}.
This severely limits data throughput, and human subjectivity is likely to introduce systematic bias.
Indeed, it is expected that scoring will be performed differently by each expert, making results difficult to reproduce independently.
Often, two experts score the same data, in order to ensure satisfying agreement.
Although, manual scorers are generally reported as being very consensual with
each other\cite{costa-miserachs_automated_2003,sen_comparative_2014}, it can be
argued that experts most likely work in the same laboratory and trained one another, or were trained by the same third person.
In this context, agreement between experts does not account for the variability between communities of researchers, and cannot be used to assess reproducibility.

In order to overcome both speed and subjectivity limitations, efforts have long
been directed towards automation of sleep
scoring\cite{chouvet_automatic_1980, haustein_automatic_1986}.
However, little adoption has occurred and very few available implementations, in the form of software that biologists could use, have been developed.
Typically, two different approaches to classification have been followed: unsupervised or supervised learning.

Unsupervised learning \cite{langkvist_sleep_2012,sunagawa_faster:_2013} has the
advantage of making no assumption about the nature of the different vigilance states, and how they should be defined.
Therefore, this approach can lead to the discovery of truly new states.
One issue is that the choice of the variables used for clustering is very
critical.
Often, variables such as frequency domain variables are in fact chosen in order
to generate clusters that will match human defined clusters.
In addition, unsupervised methods may lack robustness\cite{sunagawa_faster:_2013} in so far as they
cannot easily include covariates explaining, for instance, variability between recording equipments.

Another approach is to assume that human annotations are, although imperfect, biologically relevant and generally consistent,
 and to therefore use supervised learning
 techniques\cite{crisler_sleep-stage_2008,ventouras_performance_2012,doroshenkov_classification_2007,pan_transition-constrained_2012,sen_comparative_2014}.
  Of course, if human decisions were biased, such a method may suffer from the same bias.
However, a vast corpus of experimental work has provided hypothesis about function of these states which tends to validate the actual `existence' of these discrete vigilance states\cite{diekelmann_memory_2010}.
Building a classifier that would produce a consensual prediction of vigilance states could be seen as an attempt to formalised and rationalise the definition of such states.
This would improve future research without denying decades of sleep neurobiology.

Many supervised learning techniques such as from
\glspl{svm}\cite{crisler_sleep-stage_2008},
\glspl{ann}\cite{ventouras_performance_2012},
to
\glspl{hmm}\cite{doroshenkov_classification_2007,pan_transition-constrained_2012} have been implemented.
In general, the first step is to compute features on consecutive segments of annotated electrophysiological signals known as epochs.
Then, the relation between the response variable (annotation) and the independent variables (features) can be modelled.
Either epochs are considered to be independent from one another or to be time-dependent structures that are explicitly modelled (\eg{} using \glspl{hmm}).
Time-aware modelling has the advantage of accounting for the interdependence of consecutive epochs (see fig.~\ref{fig:sleep_description}B).
However, it generally does not model non-linear relationships between large numbers of predictors as well as classical classifiers.

Recently, promising results were obtained for automatic scoring of human sleep stages by performing an exhaustive
feature extraction, including variables resulting from discrete wavelet
decomposition\cite{sen_comparative_2014}.
Then, the authors compared several classifiers and found that random
forests\cite{breiman_random_2001} were, overall, the most accurate predictors.

The study herein bases itself on these promising results by computing an even larger number of features.
An important addition was the computation of interval
features (\ie{} time aware)
\cite{dietterich_machine_2002,rodriguez_support_2005,deng_time_2013} which significantly improved
accuracy. 
Indeed, since manual scorers usually use contextual information to make a decision concerning a given state (personal communications),
it was highly relevant to build a time-aware predictor.
Furthermore, as opposed to previous work, stratified cross-validation\cite{ding_querying_2008} procedures and
comparisons of resulting sleep patterns were performed.
These improvements altogether contributed to achieving a very satisfying overall accuracy of 92\% 
compared to manual scoring.


In order to pave the way for the implementation of a ubiquitous sleep scoring software,
\pr, a new \py{} package was also build to facilitate efficient feature extraction.
This new package is here demonstrated to perform significantly better than \pyeeg{}\cite{bao_pyeeg:_2011}, an alternative implementation.
In addition, inconsistencies in the implementations of some algorithms in \pyeeg{} were corrected, and many new features were developed.
