

\begin {table}[H]
\begin{center}
\label{table:benchmark}


\caption{\ctit{Performance improvements over \texttt{PyEEG}.}
 In order to improve efficiency, modification of the algorithms implemented in \texttt{PyEEG} were caried out.
Several mathematical inconsistancies were also discovered and corrected.
The table compares how long, on average, each algorithm would take, for a random sequence of length $1280$.
In addition, it represents how many added points would lead to a tenfold runtime increase.
For the tested range ($n \in [1280;7680] $), all algorithms add an 
exponential time complexity ($10^{O(n)}$), $R^2 > 0.95$, for all.
Significance levels were assessed by fitting linear models, on the log-transformed runtimes.
t-tests on the effect of the implementation (\ie{} whether \pr{} or \texttt{PyEEG} was used) on the intercept (at 1280, as oposed to 0) were used to compute $p-values$ at $n=1280$.
$p-values$ for the differences in scaling were obtained through t-tests on the interaction between $n$, and the implementation.
\emph{No} multiple test correction was applied to account for testing significances on different algorithms.
\textbf{1}: whether the original implementation was corrected in order to match mathematical definition.
Each alteration is mathematically justified in the section \texttt{pyrem.univariate} of the \pr{} documentation (see appendix).
}
\small
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  &  & \multicolumn{2}{|c|}{\texttt{PyEEG}} & \multicolumn{2}{|c|}{\pr} & \\
 \hline
 
  algorithm & function & \specialcell{$t$(ms) for \\$n = 1280$} & \specialcell{$n$ for $\times 10$\\increase} & \specialcell{$t$(ms) for \\$n = 1280$} & \specialcell{$n$ for $\times 10$\\ increase} & fix$^1$\\
  
  \hline
  \hline
\specialcell{Approximate\\Entropy} & \texttt{ap\_ent} & 									9970 & 4288 & $487^{***}$ & $3478^{***}$ & No\\
\hline
Fisher Information & \texttt{fisher\_info} & 								3.24 & 8673 & $0.121^{***}$ & $12427^{***}$ & No\\
\hline
\specialcell{Higuchi\\Fractal Dimension} & \texttt{hfd} &	 				11.7 & 8833 & $1.39^{***}$ & $28329^{***}$ & Yes\\
\hline
Hjorth parameters & \texttt{hjorth} & 										5.14 & 8633 & $0.088^{***}$ & $36354^{***}$ & Yes\\
\hline
\specialcell{Petrosian\\Fractal Dimension} & \texttt{pfd} & 				2.66 & 8606 & 2.65 & 8579 & Yes\\
\hline
Sample Entropy & \texttt{samp\_ent} & 										8305 & 4276 & $188^{***}$ & $5483^{***}$ & No\\
\hline
Spectral Entropy & \texttt{spectral\_entropy} & 								0.309 & 11459 & $0.227^{***}$ & $22133^{***}$ & Yes\\
\hline
\specialcell[l]{Singular Value \\Decomposition\\ entropy} & \texttt{svd\_ent} & 	3.25 & 8663 & $0.113^{***}$ & $11774^{**}$ & Yes\\
 \hline


  %~ \hline
   %~ a & f & ff & f& f&&
  %~ \hline
  %~ 3700 & octal \\ \cline{2-2}
  %~ 11111000000 & binary \\
  %~ \hline 
  %~ 1984 & decimal \\
  %~ \hline
\end{tabular}
\end{center}
\end{table}
